---
title: "Urchin Populations in Santa Barbara Reefs"
author: "Margaret Hammond"
date: "10/26/2021"
output: html_document
---

## Overview
Sea urchin populations in 5 different reefs in Santa Barbara: Naples Reef, Arroyo Quemado Reef, Carpinteria Reef, Mohawk Reef, and Isla Vista Reef. This data holds information on the size frequency distribution of red and purple sea urchins from 2008 - 2020. The data is analyzed to determine the difference of size between red and purple sea urchins, and if the sea urchin population is decreasing among the sites. 

## Citation
Data: Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b. 


```{r setup, include = TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(kableExtra)
library(here)
library(janitor)
library(lubridate)
```

## Read in Data

```{r}
urchins <- read_csv(here("data", "urchins.csv"))

urchins_clean <- urchins %>% clean_names()
```

## Data Visualization

Histogram to compare the sizes of red and purple sea urchin populations in Santa Barbara reefs. Bins lowered to 10 per urchin type for cleaner organization.

```{r}

ggplot(data = urchins_clean, aes(x = size)) +
  geom_histogram(bins = 10,
                 aes(fill = common_name)) +
  scale_fill_manual(values = c("purple", "red")) +
  facet_wrap(~ common_name) +
  labs(x = "Size (cm)",
       y = "Count",
       title = "Size of Sea Urchin (cm) by Common Name",
       fill = "Common Name",
       caption = "Figure 1. Diameter (cm) of Red and Purple 
       Sea Urchin populations within Santa Barbara reef habitats.") +
  theme_minimal()
```


## Statistics Table

Summary table that shows the statistics for purple and red urchins. Includes mean, median, standard deviation, and sample size. This is to understand the data statistics and compare the sizes between the two urchin populations.


```{r}
urchin_stats <- urchins_clean %>% 
  group_by(common_name) %>% 
  summarize(mean_size_cm = mean(size, na.rm = TRUE), 
            median_size_cm = median(size), 
            sd_size_cm = sd(size),
            sample_size = n())

urchin_stats %>% 
  kable(col.names = c("Common Name",
                      "Mean size (cm)", 
                      "Median size (cm)",
                      "Standard deviation (cm)",
                      "Sample size (n)"),
        caption = "Figure 2. Sea Urchin Size Statistics Summary by Common Name") %>% 
  kable_styling(full_width = TRUE) 

```


## Data Analysis
From the histogram and table created above, we are able to assess the similarities and differences between the mean diameter size of purple and red sea urchins. The data shows that the red sea urchin, on average, is larger than the purple sea urchin. As shown in the table, the red sea urchin is about 2.3 inches larger in diameter. The red sea urchin is about 51% bigger than the purple sea urchin. The red sea urchin also has a larger range and a more normal distribution than the purple sea urchin. 


## Purple Urchin Data Creation
Create a separate data set of purple urchins. In this data set, date is cleaned up to be in proper R format, and a new column `year` is created. Column `site_full` created to give the full names of the site collection spots. 

```{r}
purple_urchins <- urchins_clean %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(date = mdy(date)) %>% 
  mutate(year = year(date)) %>% 
  mutate(site_full = case_when(
    site == "NAPL" ~ "Naples Reef",
    site == "AQUE" ~ "Arroyo Quemado Reef",
    site == "CARP" ~ "Carpinteria Reef",
    site == "MOHK" ~ "Mohawk Reef",
    site == "IVEE" ~ "Isla Vista Reef"
  ))
  
```


## Table Set Up

Creating the code for a table to show the total number of purple urchins observed by site and year. 

```{r}
purple_urchins_table <- purple_urchins %>% 
  count(site_full, year) 

```


## Graph of Purple Urchins

Make a finalized graph of counts of purple urchins recorded at each site over time for the years included in the study. 

```{r}

#ggplot(data = purple_urchins_table, aes(x = year, y = n,)) +
#  geom_line(aes(color = site_full)) +
#  theme(text = element_text(
 #   size = 9)) +
 #     theme(axis.text.x = element_text(angle = 45)) +
 # labs(x = "Year",
 #      y = "Amount of Urchins",
 #      title = "Purple Sea Urchin Population by Site",
 #      color = "Site",
 #      caption = "Figure 2. Purple Sea Urchin Populations Over Time at 5 sites") 

ggplot(data = purple_urchins_table, aes(x = year, y = n)) +
 geom_bar(aes(fill = site_full), stat = "identity", color = "black") +
  facet_wrap(~ site_full, scale = "free") +
  theme(text = element_text(
    size = 9)) +
      theme(axis.text.x = element_text(angle = 45)) +
  labs(x = "Year",
       y = "Amount of Urchins",
       title = "Purple Sea Urchin Population by Site",
       fill = "Site",
       caption = "Figure 2. Purple Sea Urchin Population Over Time at 5 Reef Sites in Santa Barbara") 

```


## Data Analysis

This graph shows a decline of purple sea urchin populations at all sites during the final years of observation. This may not be reflective of the total purple sea urchin population, because the sea urchins may be moving to other, unrecorded sites. We cannot make conclusions about the total population decline without understanding possible habitat changes and population dynamics. 


## End of Task 3



